Pollution, agricultural productivity, and development: Evidence from coal in plants in India

### 2023-02-11 presentation/slides/slides.Rmd
updating tables

library(sf)
library(sp)
library(tidyverse)
library(gganimate)
library(transformr)
library(gifski)
library(ggthemes)
library(scales)
library(RColorBrewer)
library(kableExtra)
library(ggpubr)
library(rticles)
library(stats)
library(broom)
library(extrafont)
library(fixest)
library(modelsummary)
library(openxlsx)
library(raster)
library(lubridate)
loadfonts()  # Must be run once in each session (markdown is a new "session" each time you render/knit)

@@

@@

@@

ggplot() +
    geom_sf(data = districts, color = "gray85", fill = "gray80") +
    geom_sf(data = villageswind, 
            aes(fill = indicator),
            alpha = .5) +
    # Here comes the gganimate code
    transition_manual(
                      frames = date,
                      cumulative = FALSE
                      ) +
    labs(title = "date: {current_frame}") +
    theme(legend.position = "none")

@@

# Table
colnames(plantresultstable) <- c("1991", "2001", "2001", "2011")
kable(
      plantresultstable, caption = "Local characteristics and the construction of coal plants",
      align = "cccc", booktabs = TRUE, linesep = "", escape = FALSE
      ) %>%
  footnote(
           "Note: Robust standard errors are in parentheses. The outcome in the first column is whether the village is within 30km of a coal plant in 1991. The second column is whether a village in 1991 will have a coal plant in 2001, conditional on not having one in 1991. The last two columns are similarly defined, except using 2001 and 2011 as the years. * p<0.10 ** p<0.05 *** p<0.01",
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE,
           escape = FALSE
           ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:5),width = "2cm") %>%
  row_spec(9, hline_after = TRUE) %>%
  add_header_above(c(" ", "1991 census" = 2, "2001 census" = 2)) %>%
  kable_classic_2()

@@

## Validating wind direction

@@

# Table
colnames(pollutiontable) <- c("(1)", "(2)", "(3)", "(4)")
kable(
      pollutiontable, caption = "Wind direction and particulate matter",
      align = "cccc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "4cm") %>%
  column_spec(c(2:5),width = "2cm") %>%
  row_spec(c(3), bold = TRUE) %>%
  row_spec(6, hline_after = TRUE) %>%
  add_header_above(c(" ", "1998-2015" = 2, "2002-2013" = 2)) %>%
  kable_classic_2()

@@

# Table
colnames(yieldtable) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
kable(
      yieldtable, caption = "Wind direction and agricultural productivity",
      align = "ccccc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:6),width = "2cm") %>%
  row_spec(7, bold = TRUE) %>%
  row_spec(11, hline_after = TRUE) %>%
  add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
  kable_classic_2()

@@

@@

# Table
colnames(yieldtablepm) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
kable(
      yieldtablepm, caption = "Particulate matter (PM 2.5) and agricultural productivity",
      align = "ccccc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:6),width = "2cm") %>%
  row_spec(7, bold = TRUE) %>%
  row_spec(11, hline_after = TRUE) %>%
  add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
  kable_classic_2()

@@

@@

# Table
colnames(yieldtabletwo) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
kable(
      yieldtabletwo, caption = "Pollution and agricultural productivity, IV estimates",
      align = "cccc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:6),width = "2cm") %>%
  row_spec(c(7,13), bold = TRUE) %>%
  row_spec(c(11:12), hline_after = TRUE) %>%
  add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
  kable_classic_2()

@@

# Table
labortableold <- labortableold[,-1]
colnames(labortableold) <- c("all", "self", "wage", "farm", "non-farm")
kable(
      labortableold, caption = "Wind direction and labor allocation, heterogeneity by age",
      align = "ccccc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the district level. Control variables include female, age, age squared, and (years of) education. * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "4cm") %>%
  column_spec(c(2:6),width = "1.5cm") %>%
  row_spec(c(10), hline_after = TRUE) %>%
  row_spec(c(6,9), bold = TRUE) %>%
  kable_classic_2()

@@

@@

# Table
colnames(ntltable) <- c("(1)", "(2)")
kable(
      ntltable, caption = "Wind direction and nightlight growth",
      align = "cc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:3),width = "2cm") %>%
  row_spec(c(9), hline_after = TRUE) %>%
  row_spec(c(5,8), bold = TRUE) %>%
  kable_classic_2()

@@

>>>>

library(sf)
library(sp)
library(tidyverse)
library(gganimate)
library(transformr)
library(gifski)
library(ggthemes)
library(scales)
library(RColorBrewer)
library(kableExtra)
library(ggpubr)
library(rticles)
library(stats)
library(broom)
library(extrafont)
library(fixest)
library(modelsummary)
library(openxlsx)
library(raster)
library(lubridate)
library(icons)
library(png)
loadfonts()  # Must be run once in each session (markdown is a new "session" each time you render/knit)

@@

```{css, echo=F}
    /* Table width = 100% max-width */

    <!-- .remark-slide table{ -->
    <!--     width: 100%; -->
    <!-- } -->

    /* Change the background color to white for shaded rows (even rows) */

    .remark-slide thead, .remark-slide tr:nth-child(n) {
        background-color: white;
    }

    tfoot {
      font-size: 80%;
    }
```

## Coal and energy worldwide

- Coal one of most common energy sources worldwide
  - Reached an all-time high in 2021 <small><small>(www.iea.org/reports/coal-fired-electricity)</small></small>

![]("../../../pollution_development/draft/tables/energy.png")

## Coal and energy in India

![]("../../../pollution_development/draft/tables/energy.png")

## Coal and energy in India

- From 1990 to 2010: coal generation from 42.4 gigawatts to more than 100
- Obvious benefits to more power <small><small>(Dinkelman 2011; Kline and Moretti 2014; Rud 2012; Van de Walle et al. 2017)</small></small>

## Downsides, too

- But power generation can have downsides, too
  - Especially true near plants: pollution a major concern
- Climate change and pollutants like SO2, mercury, etc. <small><small>(www.eia.gov/energyexplained/coal/coal-and-the-environment.php)</small></small>
- Long literature on negative effects of pollution 
  - Health <small><small>(Arceo, Hanna, and Oliva 2016; Brunekreef and Holgate 2002; Deryugina et al. 2019; Kampa and Castanas 2008; Pope III and Dockery 2006)</small></small>
  - Productivity/labor allocation <small><small>(Chang et al. 2019; Ebenstein, Lavy, and Roth 2016; Hanna and Oliva 2015; He, Liu, and Salvo 2019; Wen and Burke 2022)</small></small>

## This paper

- Effects of pollution from coal plants on agricultural productivity in India
  - Previous evidence from gold mines (Aragon and Rud 2016) and workers in California (Graff Zivin and Neidell 2012)
- Higher exposure to pollution leads to lower agricultural productivity
  - Exposure defined as wind direction from coal plants towards villages
  - One additional day of exposure per season: decrease of 0.3 percent (yields)

## This paper

- Adults decrease labor allocation
  - Driven by large decrease in older half of sample; younger half increases allocation to farm labor
- More pollution in previous year also leads to lower nightlight growth
  - 25th to 75th percentile of exposure leads to decrease in nightlights of 0.7 percent

## Data sources

```{r data, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

data <- matrix(NA, nrow = 8, ncol = 3)
rownames(data) <- c("shapefile", "coal plants", "wind", "pollution", "agriculture", "nightlights", "weather", "labor")
colnames(data) <- c("source", "geographic coverage", "temporal coverage")
data[1,1] <- "Asher et al. (2021)"
data[2,1] <- "Global Energy Monitor"
data[3,1] <- "NCAR"
data[4,1] <- "Hammer et al. (2020)"
data[5,1] <- "Angopadhyay et al. (2022)"
data[6,1] <- "Asher et al. (2021)"
data[7,1] <- "TerraClimate"
data[8,1] <- "National Sample Survey (NSS)"
data[1,2] <- "India"
data[2,2] <- "global"
data[3,2] <- "global"
data[4,2] <- "global"
data[5,2] <- "India"
data[6,2] <- "India"
data[7,2] <- "global"
data[8,2] <- "India"
data[1,3] <- " "
data[2,3] <- "yearly"
data[3,3] <- "daily"
data[4,3] <- "monthly"
data[5,3] <- "two seasons/year"
data[6,3] <- "yearly"
data[7,3] <- "monthly"
data[8,3] <- "daily interviews"

# Table
kable(
      data,
      align = "ccccc", booktabs = TRUE, linesep = "", escape = TRUE
      ) %>%
  footnote(
           general = "Global Energy Monitor: globalenergymonitor.org/projects/global-coal-plant-tracker. NCAR: climatedataguide.ucar.edu/. TerraClimate: www.climatologylab.org/terraclimate.html",
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "2cm") %>%
  column_spec(2,width = "4.5cm") %>%
  column_spec(c(3:4),width = "3.5cm") %>%
  kable_styling(latex_options = "HOLD_position") %>%
  kable_classic_2()

## General idea
- Coal plants open in a given year
- Plot wind direction each day
  - Exposure means wind blowing from coal plant to village (within 30km)
  - Aggregate up to desired level (e.g. season, month, week)

@@

```{r windbase, echo = FALSE, message = FALSE, warning = FALSE,}

ggplot() +
    geom_sf(data = districts, color = "gray85", fill = "gray80") +
    geom_sf(data = villageswind %>% filter(date=="2010-01-01"),
            alpha = .5) +
    labs(title = "date: 2010-01-01") +
    theme(legend.position = "none")

@@

ggplot() +
    geom_sf(data = villageswind, 
            aes(fill = indicator),
            alpha = .5) +
    # Here comes the gganimate code
    transition_manual(
                      frames = date,
                      cumulative = FALSE
                      ) +
    labs(title = "date: {current_frame}") +
    theme(legend.position = "none")

@@

# Table
colnames(plantresultstable) <- c("1991", "2001", "2001", "2011")
kable(
      plantresultstable,
      align = "cccc", booktabs = TRUE, linesep = "", escape = FALSE
      ) %>%
  footnote(
           "Note: Robust standard errors are in parentheses. The outcome in the first column is whether the village is within 30km of a coal plant in 1991. The second column is whether a village in 1991 will have a coal plant in 2001, conditional on not having one in 1991. The last two columns are similarly defined, except using 2001 and 2011 as the years. * p<0.10 ** p<0.05 *** p<0.01",
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE,
           escape = FALSE
           ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:5),width = "2cm") %>%
  row_spec(9, hline_after = TRUE) %>%
  add_header_above(c(" ", "1991 census" = 2, "2001 census" = 2)) %>%
  kable_classic_2()

@@

## Validating wind direction - particulate matter and exposure

@@

# Table
colnames(pollutiontable) <- c("(1)", "(2)", "(3)", "(4)")
kable(
      pollutiontable,
      align = "cccc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "4cm") %>%
  column_spec(c(2:5),width = "2cm") %>%
  row_spec(c(3), bold = TRUE) %>%
  row_spec(6, hline_after = TRUE) %>%
  add_header_above(c(" ", "1998-2015" = 2, "2002-2013" = 2)) %>%
  kable_classic_2()

@@

# Table
colnames(yieldtable) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
kable(
      yieldtable,
      align = "ccccc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:6),width = "2cm") %>%
  row_spec(7, bold = TRUE) %>%
  row_spec(11, hline_after = TRUE) %>%
  add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
  kable_classic_2()

@@

- Likely endogenous
  - Upward biased?

@@

# Table
colnames(yieldtablepm) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
kable(
      yieldtablepm[-c(3:6),],
      align = "ccccc", booktabs = TRUE, linesep = ""
      ) %>%
  # footnote(
  #          "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
  #          general_title = "",
  #          threeparttable = TRUE,
  #          footnote_as_chunk = TRUE
  #          ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:6),width = "2cm") %>%
  row_spec(3, bold = TRUE) %>%
  row_spec(7, hline_after = TRUE) %>%
  add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
  kable_classic_2()

@@

- We are really interested in the effect of _pollution_, not wind
- Use wind as instrument? Assumes...
  - Conditional on fixed effects and weather...
  - Wind only affects agriculture through pollution
- Reasonable?

@@

# Table
colnames(yieldtabletwo) <- c("(1)", "(2)", "(3)", "(4)", "(5)")
kable(
      yieldtabletwo[-c(3:6, 12, 16:19),],
      align = "cccc", booktabs = TRUE, linesep = ""
      ) %>%
  # footnote(
  #          "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
  #          general_title = "",
  #          threeparttable = TRUE,
  #          footnote_as_chunk = TRUE
  #          ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:6),width = "2cm") %>%
  row_spec(c(3,8), bold = TRUE) %>%
  row_spec(c(2), hline_after = TRUE) %>%
  add_header_above(c(" ", "all" = 3, "monsoon" = 1, "winter" = 1)) %>%
  kable_classic_2()

## Robustness checks

- Leads of wind direction
- Only post coal plant opening

## Taking stock

- We see clear negative effects of pollution on productivity
  - Absolute deviation of exposure is approx. 8 days, meaning 2-3 percent change in productivity
  - Using AD of PM, it's above 20 percent change
- Key question: what is the driver?
  - Land?
  - Labor?
- Hard to look explicitly at land, so let's look at labor

## National Sample Survey (NSS)

```{r nss, echo = FALSE, message = FALSE, warning = FALSE,}

laborplot

## Problem: only have district identifiers, so aggregate up

```{r aggregation, echo = FALSE, message = FALSE, warning = FALSE,}

# Sets wd to folder this script is in so we can create relative paths instead of absolute paths
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
# We want to get the wd up to the main folder
# Need to go up three levels
setwd('..')
setwd('..')
setwd('..')

# district
districts <- read_sf(paste0("data/spatial/districts/districts.shp"))
districts <- districts %>% 
                dplyr::select(state = ST_CEN_CD, district = DT_CEN_CD, state_name = ST_NM, district_name = DISTRICT)
districts <- districts %>% filter(state=="23" & district=="07")
# Villages
villages <- read_sf(paste0("data/spatial/villages_overlap/villages_overlap.shp"))
villages <- villages %>% mutate(shrid = paste0(pc11_s_id, "-", pc11_tv_id)) %>%
                          dplyr::select(shrid, state = ST_CEN_CD, district = DT_CEN_CD, state_name = ST_NM, district_name = DISTRICT)
villages <- villages %>% filter(state=="23" & district=="07")
wind <- read_csv(paste0("data/clean/wind_ntl/days/date_2010-1-1.csv")) %>% as_tibble()
wind$days_sum <- apply(wind[,2:8], 1, FUN = "sum")
wind <- wind %>% dplyr::select(shrid, days_sum)
villages <- villages %>% left_join(wind, by = "shrid")
villages <- villages %>% st_set_crs(st_crs(districts))

ggplot() + 
  theme_map() +
  geom_sf(data = districts, fill = NA) +
  geom_sf(data = villages %>% filter(is.na(days_sum)==F), aes(fill = days_sum), lwd = NA) +
  scale_fill_viridis_c("exposure (days)", option = "viridis") +
  labs(x = "lon",
       y = "lat") +
  theme(legend.position = c(0.5, 0.05),
        legend.direction = "horizontal",
        legend.key.width = unit(0.3, 'cm'),
        legend.text = element_text(size = 8),
        legend.background=element_blank() )

@@

# Table
labortableold <- labortableold[,-1]
colnames(labortableold) <- c("all", "self", "wage", "farm", "non-farm")
kable(
      labortableold,
      align = "ccccc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the district level. Control variables include female, age, age squared, and (years of) education. * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "4cm") %>%
  column_spec(c(2:6),width = "1.5cm") %>%
  row_spec(c(10), hline_after = TRUE) %>%
  row_spec(c(6,9), bold = TRUE) %>%
  kable_classic_2()

@@

## Large negative effects on older half

- Older half of sample decreases overall labor allocation
  - Driven by self-employment and nonfarm employment
- Younger half actually increases allocation!
- Consistent with health effects only on older adults, though not conclusive

## Effects on aggregate growth?

- We know pollution affects productivity and labor allocation
- Might it also affect growth?
- Proxy for growth with nightlights
  - Look at previous year's exposure and current year's nightlights
  - "Growth" of nightlights
  - Village FE

@@

# Table
colnames(ntltable) <- c("(1)", "(2)")
kable(
      ntltable,
      align = "cc", booktabs = TRUE, linesep = ""
      ) %>%
  footnote(
           "Note: Standard errors are in parentheses and are clustered at the village level. * p<0.10 ** p<0.05 *** p<0.01", 
           general_title = "",
           threeparttable = TRUE,
           footnote_as_chunk = TRUE
           ) %>%
  column_spec(1, width = "3cm") %>%
  column_spec(c(2:3),width = "2cm") %>%
  row_spec(c(9), hline_after = TRUE) %>%
  row_spec(c(5,8), bold = TRUE) %>%
  kable_classic_2()

@@

- Large growth in coal power in India
  - Other countries, as well
- Increase in electricity, but downsides as well
  - Pollution!
- We already have plenty of evidence on effects of pollution on health

- Paper presents evidence that villages downwind from coal plants suffer in terms of agricultural productivity
  - Identification relies on within-village changes across seasons
- Wind exposure fluctuations lead to changes in agricultural productivity of around 2.4 percent based on mean absolute deviation
  - Instrumenting for particulate matter, effects are even larger (more than 20 percent)
- Clear changes in labor allocation
  - Oldest half of sample
- Slower growth of nightlights

- Overall cost/benefit way beyond scope of paper
- Instead, results point to importance of the location of pollution sources
  - Downwind villages suffer
- 30km radius indicates relatively long-range effects
